---
title: "Hw04"
output: github_document 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework 4

## Part 1, Prompt 1: Jian's tidyr cheatsheet

Sometimes our data could be either too messy or too genralized that we cannot analyze the properly. Sometimes, we want to make certain plots but the varibles is not directly presented that R could understand. Therefore, we need to reshape our data frame so that we could have the information we needed to poeperly analyze the data. 
**Gather** and **Spread** are two useful tools in tidyr in helping us reshape our data. Both functions will create new columns that either combines variables into one (make the data frame longer) or split one varible into many varibles(make the data frame wider). 

## Format 
The format of gather and spread is:
data.frame %>% gather/spread(key,value,...)

**Key** will be the header(s) for the new columns that include previous variable headers
**Value** will be the header(s) for the new columns that include previous varibale observations
**...** will be the reshaping criterias, what include and what not include. 

I will use gapminder data to demonstrate how gather and spread could be used. 

Let's say we want to know the total GDP of Canada and China. 

```{r}
library(gapminder)
library(tidyverse)
```

```{r}
example1 <- gapminder %>% 
  filter(country %in% c("China","Canada"))%>%
  select(gdpPercap,pop,year,country)%>%
  mutate(gdp_billions=gdpPercap*pop/1000000000)%>%
  select(gdp_billions,year,country)

knitr::kable(example1)
```





This would be how the data table looks like by default. However, if I want to compare the GDP between the two countries at each individual years, it would be much easier to reshape the lengthy table according to years by using the spread function. 

```{r}
example1spread <- spread(example1,year,gdp_billions)
knitr::kable(example1spread)
```


We can also categorize the columns based on country 

```{r}
example2spread <- spread(example1,country,gdp_billions)
knitr::kable(example2spread)
```

Now, it is much easier to compare the data. 

The gather function makes the table longer by combining variables.For example, if I want to make GDP of China and Canada into one column and specify the country and one column specify the GDP, I could use gather.

```{r}
example1gather <- gather(example2spread,country,gdp_billions,Canada,China)
knitr::kable(example1gather)
```


We could also merge two variable into one column by **unite**

```{r}
example1unite <- unite(example1,"Country_year",c(country,year),sep = ",")
knitr::kable(example1unite)
```


On the other hand, we could also seperate one column into two columns by **seperate**

```{r}
example1seperate <- separate(example1unite,Country_year,c("Country","year"),sep = ",")
knitr::kable(example1seperate)

```


## Part 2, Prompt 2: Jian's joinging cheatsheet

In this section, I have created two csv file about dogs to be my example. 

First, I will load the data. 

```{r}
dog <- read_csv("~/Desktop/Stat 545/dog1.csv")
breed <- read_csv("~/Desktop/Stat 545/dog2.csv")
knitr::kable(dog)
knitr::kable(breed)
```

The mission is combine the two data frames into one in various styles by using the dplyr join function. 


## 1. Inner_join 

Inner join is useful in combining the two data frame together. The format of inner join is inner_join(x,y). In this case, x will be the main data frame and y will be the adding frame. In other words, the format of X will not change but y will be added into x according to x's format. 

```{r}
dogsample1 <- inner_join(dog,breed)
knitr::kable(dogsample1)
```
 

In this case, x is dog and y is breed data. The arrangement of dog data did not change but the breed data is added into dog data according to the common column "breed group". However, since there is no "hound" in the breed file, we have lost beagle in the joined frame. Let's try what will the table looks like if we use breed as x and dog to be y. 

```{r}
dogsample2 <- inner_join(breed,dog)
knitr::kable(dogsample2)
```


Now, we see that data from y will be added according to x's format. 


## 2. Semi_join 

Next, there is semi_join. Semi join differs from inner_join as not column besides the common column will be included in the final result. For example 

```{r}
dogsample3 <- semi_join(dog,breed)
knitr::kable(dogsample3)
```


## 3. Left_join 

Left_join includes all the matching pairs as well as the non-paired. In other word, all the data will be included in the joined version. 

```{r}
dogsample4 <- left_join(dog,breed)
knitr::kable(dogsample4)
```

In this case, left_join provides a more thorough join. 

## 4. Anti_join 

Anti_join includes non-paired join. We can use this function to filter for the data that present only in one of the data frame. 

```{r}
dogsample5 <- anti_join(dog,breed)
knitr::kable(dogsample5)
```

By utilizing these join functions, we can glue together the information we actually will be focus on and filter out the unneccessary data. 







